\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis

### 2020-10-29 manuscript.Rnw
A start on the coding framework section

%% other packages
\usepackage{algorithm}

@@

%% - \Abstract{} almost as usual
\Abstract{
Health economic models simulate the costs and effects of health technologies for use in health technology assessment (HTA) to inform efficient use of scarce resources. Models have historically been developed using spreadsheet software (e.g., Microsoft \proglang{Excel}) and while use of \proglang{R} is growing, general purpose modeling software is still limited. \pkg{hesim} helps fill this gap by facilitating parameterization, simulation, and analysis of economic models in an integrated manner. Supported model types include cohort discrete time state transition models (cDTSTMs), individual continuous time state transition models (iCTSTMs), and partitioned survival models (PSMs), encompassing Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes. A modular design based on \pkg{R6} and \proglang{S3} classes allows users to combine separate submodels for disease progression, costs, and quality-adjusted life years (QALYs) in a flexible way. Probabilistic sensitivity analysis (PSA) is used to propagate uncertainty in model parameters to model outputs. Simulation code is written in \proglang{C++} so complex simulations such as those combining PSA and individual simulation can be run much more quickly than previously possible. Decision analysis within a cost-effectiveness framework is performed using simulated costs and QALYs from a PSA.
}

%% - \Keywords{} with LaTeX markup, at least one required
%% - \Plainkeywords{} without LaTeX markup (if necessary)
%% - Should be comma-separated and in sentence case.
\Keywords{health economic evaluation, cost-effectiveness analysis, simulation, modeling, \proglang{R}}
\Plainkeywords{health economic evaluation, cost-effectiveness analysis, simulation, modeling, R}

@@

\subsection{Partitioned survival models} \label{sec:PSMs}
PSMs are conceptually similar to state transition models in that they are characterized by mutually exclusive health states. They differ, however, in that state probabilities are not computed via matrix multiplication or individual simulation, but from a set of non-mutually exclusive survival curves \citep{glasziou1990quality, woods2018nice}. Each survival curve represents time to transitioning to that state or to a more severe health state.

@@

%% -- Framework ----------------------------------------------------------------
\section{Framework} \label{sec:framework}
As shown in Figure~\ref{fig:econ-eval-process-hesim}, a typical analysis proceeds in a 3 steps:

@@

>>>>

%% other packages
\usepackage{algorithm}
\usepackage{multirow}
\usepackage{caption} 
\captionsetup[table]{skip=5pt}

@@

%% - \Abstract{} almost as usual
\Abstract{
Health economic models simulate the costs and effects of health technologies for use in health technology assessment (HTA) to inform efficient use of scarce resources. Models have historically been developed using spreadsheet software (e.g., Microsoft \proglang{Excel}) and while use of \proglang{R} is growing, general purpose modeling software is still limited. \pkg{hesim} helps fill this gap by facilitating parameterization, simulation, and analysis of economic models in an integrated manner. Supported model types include cohort discrete time state transition models (cDTSTMs), individual continuous time state transition models (iCTSTMs), and partitioned survival models (PSMs), encompassing Markov (time-homogeneous and time-inhomogeneous) and semi-Markov processes. A modular design based on \pkg{R6} and \code{S3} classes allows users to combine separate submodels for disease progression, costs, and quality-adjusted life years (QALYs) in a flexible way. Probabilistic sensitivity analysis (PSA) is used to propagate uncertainty in model parameters to model outputs. Simulation code is written in \proglang{C++} so complex simulations such as those combining PSA and individual simulation can be run much more quickly than previously possible. Decision analysis within a cost-effectiveness framework is performed using simulated costs and QALYs from a PSA.
}

%% - \Keywords{} with LaTeX markup, at least one required
%% - \Plainkeywords{} without LaTeX markup (if necessary)
%% - Should be comma-separated and in sentence case.
\Keywords{health economic evaluation, cost-effectiveness analysis, simulation, multi-state models, \proglang{R}}
\Plainkeywords{health economic evaluation, cost-effectiveness analysis, simulation, multi-state models, R}

@@

\subsection{Partitioned survival models} \label{sec:PSMs}
PSMs are conceptually similar to STMs in that they are characterized by mutually exclusive health states. They differ, however, in that state probabilities are not computed via matrix multiplication or individual simulation, but from a set of non-mutually exclusive survival curves \citep{glasziou1990quality, woods2018nice}. Each survival curve represents time to transitioning to that state or to a more severe health state.

@@

%% -- Framework ----------------------------------------------------------------
\section{Framework} \label{sec:framework}
The \pkg{hesim} coding framework aims to follow three main tenets. First, it should be possible to seamlessly integrate parameter estimation with simulation. Second, modular code show allow for models to be built in a flexible manner. Finally, code should be computationally efficient.

Each economic model consists of separate statistical models of disease progression, costs, and utilities. All analyses proceed in 4 steps. The first step is to setup the model by defining the model structure, target population, and treatment strategies of interest. This is implemented in \pkg{hesim} with the \code{hesim_data} class as will be illustrated in the examples below.

The actual analysis is performed in the next three steps as shown in Figure~\ref{fig:econ-eval-process-hesim}. First, a statistical models for disease progression, costs, and utilities are parameterized using ``estimation'' datasets. Next, the statistical models are combined to construct an economic model. For a given model structure, disease progression, QALYs, and costs are simulated from ``input data'', based on the target population and treatment strategies of interest. Finally, the simulated outcomes are used to perform decision analysis, currently limited to CEA.

@@

The three economic models described in Section~\ref{sec:model-taxonomy} are implemented as the \pkg{R6} classes \code{CohortDtstm}, \code{IndivCtstm}, and \code{Psm}. Methods are provided within each class for simulating disease progression, costs, and QALYs. These methods are made possible by separate \pkg{R6} classes for disease progression, costs, and utility, which comprise the full economic model. Below, we describe this modular design in more detail.

\subsection{Parameterization} \label{sec:parameterization}
\subsubsection{Disease progression}
The statistical model used to parameterize the disease model component of an economic model varies by the type of economic model. For example, multinomial logistic regressions can be used to parameterize a cDTSTM, a set of N-1 independent survival models are used to parameterize an N-state partitioned survival model, and multi-state models can be used to parameterize an iCTSTM (Table~\ref{tbl:parameterize-disease-model}). 
\begin{table} 
\caption{Parameterization of disease models}\label{tbl:parameterize-disease-model}
\footnotesize
\begin{tabular}{l l l l}
\hline
Economic model & Statistical model & Parameter object & Model fit object\\
\hline
\multirow{2}{*}{\code{CohortDtstm}}& Custom & \code{tparams_transprobs} & \code{define_model()}\\
 & Multinomial logistic regressions & \code{params_mlogit} & \code{multinom_list}\\
\multirow{2}{*}{\code{IndivCtstm}} & Multi-state model (joint likelihood) & \code{params_surv} & \code{flexsurv::flexsurvreg}\\
& Multi-state model (transition-specific) & \code{params_surv_list} & \code{flexsurvreg_list}\\
\code{Psm} & Independent survival models & \code{param_surv_list} & \code{flexsurvreg_list}\\
\hline
\end{tabular}
\end{table}
